---
title: "Brownlows Bets"
author: "Onalee Garusinghe"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "11/20/24"    
    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
source(file="BrLw.R")
```


<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.

## Our strategy is to build a Function and the use it to answer the three questions. 

The function will be given a class_size by number of students, odds, and Number_of_times brownlow made the bet.  It will give you back how much Brownlows earnings were. We will use it to answer our three questions below. If you can, improve this function by giving default values to some of its inputs. 

# In the class and payoff as described  If you highered 40 brownlow substitute teachers to do the same thing that Brownlow did then.
```{r}
myBrLw <- numeric(1000)
for(i in 1:1000){
myBrLw[i]=BrLw()}
```



```{r}
favstats(myBrLw)
```



## what is their average earnings
  The average earnings is "r mean(myBrLw)"
  
## the standard deviation and a 95% confidence interval for their earnings?
   1000 BrLws in order to find a 95% confidence interval to see how many students in a class of 35 have the same birthday.This will be sorted by earnings from low to high in order to find a 95% confidence interval on the high and low ends.
   
<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows winnings (or losings)-->

```{r}
options(digits = 2)
teachers <- numeric(40)
for (i in 1:40){
  teachers[i]<-BrLw()}
fav_stats(teachers) -> profit
print("we can say that Mr. Brownlows is making an average of")
print(profit$mean)
print("The 95% confidence level goes from")
print(profit$mean-2*profit$sd)
print("to")
print(profit$mean+2*profit$sd)
```

```{r}
central <- favstats(teachers)
cat("Mr Brownlow is on average making $",central$mean," over the course of the experiemnt.\n")
cat("95% Conficence Interval: Low:$",central$mean - 2*central$sd,"- High $", central$mean + 2*central$sd)
```

# What are the fair odds? Use manipulation of odds in the function to tune Brownlows earnings to zero. 


<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 637 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->

```{r}
teachers <- numeric (1000)
 for (i in 1:1000){
   teachers [i] = BrLw(odds=4.38)}
 favstats(teachers)
```

  First we conducted tests on the odds with a sample of 40 teachers. Despite numerous iterations, the odds remained unstable, prompting us to expand the sample to 1000 teachers. Following this increase, we reevaluated the odds. After conducting multiple rounds of testing, we identified the set of odds with a stability factor of 4.39 as the most consistent.

```{r}
# Create a histogram
hist(teachers, 
     main = "teachers Histogram",
     xlab = "teachers",
     ylab = "Frequency",
     col = "skyblue",
     border = "black")
```

# What size of class makes the 2:1 bet fair?  Use manipulation of number of Students to tune Brownlows earnings to zero.

<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting profits!-->

```{r}
teachers <- numeric(40)
for (i in 1:40){
  teachers[i]=BrLw(numClass=28)}
favstats(teachers)
```

# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->

  In this write up we were able to discover the manipulation of odds and use trial and error methods to get our expected results. We adjusted the number of teachers from 40 to 1000 maintain a conclusion that satisfies our desired mean number. We used Google to help us create histograms for teachers earnings. Professor Brownlow would need to analyze his costs and revenue. He should consider factors such as teaching expenses, venue costs, materials, and administrative overhead. By calculating his total costs and dividing by the number of students, he can determine the per-student fee required to break even. From there, he can adjust his pricing strategy to achieve his desired profit margin, whether that's zero or otherwise.
  
  